### 2015-02-27 related-work.tex
update figure

\section{Related Work}

\paragraph{Multimethods} 
Millstein and Chambers~\cite{MS02}
describe Dubious, a simple statically typed core language including multimethods that
dispatch on the type of its arguments. They tackle a key challenge for statically typing
multimethods: ``it is possible for two modules containing arbitrary multimethods to typecheck
successfully in isolation but generate type errors when linked together.''~\cite{MS02}

% one sentence
% TC based on TR, already covered

%\paragraph{Occurrence Typing} 
%Occurrence typing~\cite{TF08,TF10} extends the type 
%system with a \emph{proposition environment} that represents 
%the information on the types of bindings down conditional branches.
%These propositions are then used to update the types associated
%with bindings in the \emph{type environment} down branches
%so binding occurrences are given different types 
%depending on the branches they appear in, and the conditionals
%that lead to that branch.

% What's diff about TC from the related work
% small summary for deisel....
% - diesel supports x
%- - calculus supports some subset of x
% we support y, which covers most of x but also foo

% eg. multiple dispatch
%     nominal vs structural

% eg. run abritrary metaprogramming over dispatch in CLOS
%  more expressive

% type systems for mm or rows
% rows vs HMap
% - no poly in HMap
% - based on subtyping
% - rows based on polymorphism

\paragraph{Record Types} 
O'Caml-style extensible record types have been the subject of extensive research 
(eg. Wand~\cite{Wan89}, Cardelli and Mitchell~\cite{CM91}, Harper and Pierce~\cite{HP91})
most inference

Dependent JavaScript~\cite{Chugh:2012:DTJ} can track similar invariants as HMaps with types
for JS objects. They must deal with mutable objects, they feature refinement types and strong updates to the heap
to track changes to objects.

Typed Lua~\cite{Maidl:2014:TLO} has \emph{table types} which track entries in a mutable Lua table.
Typed Lua changes the dynamic semantics of Lua to accommodate mutability: Typed Lua raises a runtime error
for lookups on missing keys---HMaps consider lookups on missing keys normal.

\paragraph{Optional type systems}
\begin{itemize}
  \item Reticulated Python~\cite{Vitousek14}
  \item GradualTalk
  \item Flow
\end{itemize}

Typed Racket is a statically typed sister language of Racket. It
attempts to preserve existing Racket idioms and aims to type check
existing Racket code by simply adding top level type annotations~\cite{Tob10}.

Typed Racket fully expands all macro calls before type checking~\cite{Tob10},
avoiding the complex semantics of type checking macro definitions, an ongoing research area summarised
by~\citet{Her10}. Typed Clojure also expands macros before type checking.

\paragraph{Variable-arity Polymorphism}

~\citet{STF09} invented a type system supporting variable-arity polymorphism  % doesn't fit on a line
a version of which is included in the current implementation of Typed Racket.
Their main innovation centres around \emph{dotted type variables}, which represent a heterogeneous sequence
of types. Dotted type variables allow \emph{non-uniform} variable-arity function types,
which are used to check definitions and usages of functions with non-trivial rest parameters

\paragraph{Java Interoperability in Statically Typed Languages}

Scala~\cite{OCD+} has nullable references for compatibility with Java.
Programmers must manually check for
\java{null} like in Java to avoid null-pointer exceptions.

% def checkNullOrEmpty(v:Seq[Any]):Boolean = {
%   return v.getClass
% }
% 
% checkNullOrEmpty(null) ;=> NPE
%
% null.getClass