### 2015-02-28 related-work.tex
ocaml

\section{Related Work}

\paragraph{Multimethods} 
\citet{MS02} describe Dubious, a simple statically typed core language including multimethods that
dispatch on the type of its arguments. 
In Dubious methods declare the types of arguments that they expect which corresponds to
exclusively using \clj{class} as the dispatch function in Typed Clojure.

% one sentence
% TC based on TR, already covered

%\paragraph{Occurrence Typing} 
%Occurrence typing~\cite{TF08,TF10} extends the type 
%system with a \emph{proposition environment} that represents 
%the information on the types of bindings down conditional branches.
%These propositions are then used to update the types associated
%with bindings in the \emph{type environment} down branches
%so binding occurrences are given different types 
%depending on the branches they appear in, and the conditionals
%that lead to that branch.

% What's diff about TC from the related work
% small summary for deisel....
% - diesel supports x
%- - calculus supports some subset of x
% we support y, which covers most of x but also foo

% eg. multiple dispatch
%     nominal vs structural

% eg. run abritrary metaprogramming over dispatch in CLOS
%  more expressive

% type systems for mm or rows
% rows vs HMap
% - no poly in HMap
% - based on subtyping
% - rows based on polymorphism

\paragraph{Record Types} 
OCaml-style extensible record types (eg. \citet{Wan89}, \citet{CM91}, \citet{HP91})
are defined polymorphically in terms of row variables.
HMaps in Typed Clojure are instead designed to be used with subtyping.

Dependent JavaScript~\cite{Chugh:2012:DTJ} can track similar invariants as HMaps with types
for JS objects. They must deal with mutable objects, they feature refinement types and strong updates to the heap
to track changes to objects.

Typed Lua~\cite{Maidl:2014:TLO} has \emph{table types} which track entries in a mutable Lua table.
Typed Lua changes the dynamic semantics of Lua to accommodate mutability: Typed Lua raises a runtime error
for lookups on missing keys---HMaps consider lookups on missing keys normal.

\paragraph{Optional type systems}
\begin{itemize}
  \item Reticulated Python~\cite{Vitousek14}
  \item GradualTalk
  \item Flow
\end{itemize}