### 2015-07-05 conclusion.tex
Edit conclusion.

\section{Conclusion}
\label{sec:conclusion}

Optional type systems must be designed with close attention to the
languague that they are intendend to work for.
We have therefore designed Typed Clojure, an optionally-typed version of
Clojure, with a type system that works with a wide variety of distinctive
Clojure idioms and features. Although based on the foundation of Typed
Racket's occurrence typing approach, Typed Clojure both extends the
fundamental control-flow based reasoning as well as applying it to
handle seemingly unrelated features such as multi-methods. In
addition, Typed Clojure supports crucial features such as
heterogeneous maps and Java interoperability while integrating these
features into the core type system. Not only are each of these
features important in isolation to Clojure and Typed Clojure
programmers, but they must fit together smoothly to ensure that
existing Clojure programs are easy to convert to Typed Clojure.

The result is a sound, expressive, and useful type system which, when
implemented in \coretyped with appropriate extensions, suitable for
typechecking significant amount of existing Clojure programs.
%
As a result, Typed Clojure is already successful: it is widely used in
the Clojure community among both enthusiasts and professional
programmers and receives contributions from many developers.

Our empirical analysis of existing Typed Clojure programs bears out
our design choices. Multimethods, Java interoperation, and
heterogenous maps are indeed common in both Clojure and Typed Clojure,
meaning that our type system must accomodate them. Furthermore, they
are commonly used together, and the features of each are mutually
reinforcing. Additionally, the choice to make Java's \clj{null}
explicit in the type system is validated by the many Typed Clojure
programs that use the ability to specify non-nullable types.

% Delete the following paragraphs if space is needed.

However, there is much more that Typed Clojure can provide. Most
significantly, Typed Clojure currently does not provide \emph{gradual
  typing}---interaction between typed and untyped code is unchecked and
thus unsound. We hope to explore the possibilities of using existing
mechanisms for contracts and proxies in Java and
Clojure to enable sound gradual typing for Clojure.

Additionally, the Clojure compiler is unable to use Typed Clojure's
wealth of static information to optimize programs. Addressing this
requires not only first enabling sound gradual typing, but also
integrating Typed Clojure into the Clojure tool chain more deeply, so
that its information can be passed on to the compiler.

Finally, our case study, evaluation, and broader experience indicate that Clojure
programmers still find themselves unable to use Typed Clojure on some
of their programs for lack of expressiveness. This requires continued
effort to analyze and understand the relevant features and idioms and
develop new type checking approaches.