### 2015-12-03 esop-rebuttal.txt
esop draft rebuttal

Major issues:

1) Section 5.1 should use tables instead of prose.

We will restate the data in tables.

==========================================================================

Q: What is the delta between occurrence typing in ICFP'10 and the as presented here?

A:

Q: Is TC decidable?

Q: Why can't the type system check Example 9? What restrictions are imposed on HMaps?

A: For presentational reasons, we only support a specific pattern of looking up
a known key in a union of HMaps...

Q: What specific patterns could not be supported by TC that led to CircleCI
   abandoning it?

Q: How is TC compiled/used in practice? Is it sealed from untyped namespaces?

A: core.typed

Q: "Java cannot be passed nil"?

A: This is a conservative assumption, as the paper states in the same sentence.
   In practice this is a reasonable default, as discussed in the CircleCI evaluation (5.1).

Q: Does (dissoc k) just remove its argument-entry from a map?

A: (dissoc m k) returns a new map that is like m, but without a k entry.

Q: How symbolic representations of objects in OT relate to Scala's
   path-dependent types?

A: ??

Q: Formula (3) -- do the positive/negative statements mean that d
   is of type Nat, so N_d should be read as a proposition?

A: Yes, N_d says d is a Nat, \overline{N_d} says d is not a Nat.

Q: Are Java's generic classes (in particular collections) supported
   in TC type checking algorithm?

A: No. We chose to avoid the added complexity of Java generics ..

Q: How do Clojure IDE's "incrementalize" type checking for code completion?

A: Cursive Clojure prioritizes type hints assigned to variables in the local scope
   during method name autocompletion.

Q: "Then (new[S] F "a/b") is checked against a nilable conversion..."?

A: FIXME I have no idea, looks like a bug?

Q: What is the type of (defmethod e_m e_v e_f)?

A: As shown in T-DefMethod, it returns the same type as e_m.

Q: How overlapping multi-method clauses are handled?

A: In the type system, each method is checked in isolation with respect to 
   just the current dispatch value and the dispatch function type.
   In the operational semantics, B-BetaMulti can only fire if there is an unambiguous
   match, encoded in the GM metafunction.

Q: Citation for Footnote 1?

A: FIXME

Q: What category do you consider e.g. C# 4.0, or Scala's Dynamic, or
   Dart? (WRT gradual typing)

Q: Is the additional of an Unknown or Dynamic type in a static
    type system enough to make it gradual?

Q: Is "optional" here a synonym for Bracha's "pluggable" type system?

A: Yes, because the type system does not affect the final program.
  ... But also no...

Q: One way of thinking about those
   annotations is that in fact Typed Clojure actually is a gradual type
   system as well as an optional type system, with the annotations
   effectively acting as type assumptions which will be checked
   dynamically to interact with other untyped code.

A: Typed Clojure does not check these assumptions at runtime, thus is not
   gradually typed in any sense, just optionally typed like TypeScript.

Q: What is "ann"?

A: (ann v t) 'ann'otates variable v as type t. See examples in Section 2.2.

Q: What is the "Any" type?

A: "Any" is effectively the same as (U nil Object), and has no special runtime semantics.
   This also holds in the implementation, as JVM primitives pun their boxed equivalents
   as static types, reflecting how Clojure deals with primitives.

Q: Why does 42 have the object "0"?

A: There is no known way to derive 42 from the current environment, so the object is "empty".
   More recent forumlations of occurrence typing assign the object "42" to aid symbolic
   computations.

Q: Is the complexity of multimethods worth it, given the case studies?

A: