### 2015-12-04 esop-rebuttal.txt
esop rebuttal

We thank the reviewers for their detailed feedback.  We are thrilled the
reviewers found the paper "a joy to read", "an excellent demonstration 
of occurrence typing's power", and providing a "substantial practical 
contribution".

The reviewers' primary concerns are with two issues:

1. What are the contributions beyond occurrence typing as presented
   in ICFP'10?

There are two main extensions in relation to occurrence typing.

The first is the presentation of HMap's, with support for flow checking
on sub-parts of maps. This required several extra typing rules (Fig. 10)
and subtle, but clean, extensions to the `update` rule (Fig. 12, clauses
4-7).

The second is multimethods, which is mainly contained in the T-DefMethod
typing rule. It is in the same spirit as T-If, which was present in ICFP'10,
but demonstrates occurrence typing extends gracefully to support complex 
control flow patterns.

The main takeaway is that occurrence typing as formulated in ICFP'10
serves as a valuable base to support for the nuances of a Clojure type system,
without any particular foresight from the original authors.

2. Are multimethods worth the extra complexity in the type system, given
   they are used so little in practice?

Yes.

The numbers presented in the evaluation section are possibly misleading.
Even though a low number of multimethods are defined in practice, extending
them via `defmethod` accounts for a large amount of code.

A comparable analogy to a Java code base might be assuming interfaces are
a minor feature if a small number of interfaces are defined, yet a large number
of classes extend them.

As noted in Section 5.1, these multimethods are single-dispatch on a
fixed map entry, which further gives them similar roles as Java interfaces,
which perform single dispatch on the first argument of methods.

Following are responses to specific statements or queries.

==========================================================================

Q. Section 5.1 should use tables instead of prose.

A. We will restate the data in tables.

Q: Is TC decidable? `restrict` seems problematic.

A: The calculus leaves the details of `restrict` up to implementations. core.typed
   features an ad-hoc intersection type, so `restrict` simply returns the intersection 
   of both types, thus making the occurrence typing implementation decidable.

Q: Why can't the type system check Example 9? What restrictions are imposed on HMaps?

A: For presentational reasons we only support a specific pattern of looking up
   a common, known key in a union of HMaps (notice T-GetHMap (Fig. 10) is the only rule
   supporting lookups on unions of HMaps). Example 9 is checkable in core.typed, 
   but not the calculus, and serves as a demonstration of the kinds of patterns
   that might occur in Clojure code.

Q: What specific patterns could not be supported by TC that led to CircleCI
   abandoning it?

A: We refer to the detailed blog post they published:

     http://blog.circleci.com/why-were-no-longer-using-core-typed/

Q: Are the case studies representative programs that would be good candidates for optiona
   typing?

A: We chose the CircleCI code base as it was a substantial size. The other case
   study was a randomly chosen open source project. A goal of Typed Clojure is to support
   as many Clojure idioms as possible, and we learnt valuable lessons from both projects
   in this regard.

Q: How is TC compiled/used in practice? Is it sealed from untyped namespaces?

A: TC is simply compiled to be normal Clojure code. As alluded to in Footnote 1,
   Typed Clojure is not gradually typed.

Q: "Java cannot be passed nil"?

A: This is a conservative assumption, as the paper states in the same sentence.
   In practice this is a reasonable default, as discussed in the CircleCI evaluation (5.1).

Q: Does (dissoc k) just remove its argument-entry from a map?

A: (dissoc m k) returns a new map that is like m, but without a k entry.

Q: How do symbolic representations of objects in OT relate to Scala's
   path-dependent types?

A: They are unrelated.

Q: Formula (3) -- do the positive/negative statements mean that d
   is of type Nat, so N_d should be read as a proposition?

A: Yes, N_d says d is a Nat, \overline{N_d} says d is not a Nat.

Q: Are Java's generic classes (in particular collections) supported
   in TC type checking algorithm?

A: No. TC however supports parametric polymorphism for Clojure collections and a manual subset
   of Java collections. In practice, this is sufficient for many interop situations.

Q: The simplified syntax chosen for method calls works for compilers, but
   not IDE's. How do Clojure IDE's "incrementalize" type checking for code completion?

A: This syntax is not arbitrary; (.method o) is sugar for (. o (method)) in Clojure.
   Cursive Clojure prioritizes type hints assigned to variables in the local scope
   during method name autocompletion.

Q: "Then (new[S] F "a/b") is checked against a nilable conversion..."?

A: The prose should be fixed to instead read "non-nilable".

Q: What is the type of (defmethod e_m e_v e_f)?

A: As shown in T-DefMethod, it returns the same type as e_m.

Q: How overlapping multi-method clauses are handled?

A: In the type system, each method is checked in isolation with respect to 
   just the current dispatch value and the dispatch function type.
   In the operational semantics, B-BetaMulti can only fire if there is an unambiguous
   match, encoded in the GM metafunction.

Q: Citation for Footnote 1?

A: Siek et al SNAPL'15 is a good candidate for describing gradual typing, however optional
   typing is not explicitly defined.

Q: What category do you consider e.g. C# 4.0, or Scala's Dynamic, or
   Dart? (WRT gradual typing)

A: C# 4.0 is gradually typed. Scala's Dynamic is an unrelated metaprogramming construct, 
   comparable to a statically resolved "method_missing" in Ruby.  Dart is 
   optionally typed.

Q: Is the additional of an Unknown or Dynamic type in a static
   type system enough to make it gradual?

A: A gradual type system is defined by its operational semantics, in particular
   how the runtime prevents untyped code from breaking static invariants.

Q: Is "optional" here a synonym for Bracha's "pluggable" type system?

A: They are in the same spirit.

Q: One way of thinking about those annotations is that in fact Typed Clojure 
   actually is a gradual type system as well as an optional type system, with 
   the annotations effectively acting as type assumptions which will be checked
   dynamically to interact with other untyped code.

A: Typed Clojure does not check these assumptions at runtime, thus is not
   gradually typed in any sense, just optionally typed like TypeScript.

Q: What is "ann"?

A: (ann v t) 'ann'otates variable v as type t. See examples in Section 2.2.
   We will explain this explicitly.

Q: What is the "Any" type? "Kw"?

A: "Any" is effectively the same as (U nil Object), and has no special runtime semantics.
   It is *not* a Dynamic type.
   "Kw" is the type for keywords.

Q: Why does 42 have the object "0"?

A: There is no known way to derive 42 from the current environment, so we say the 
   object is "empty".