### 2015-12-05 esop-rebuttal.txt
Edits.

We thank the reviewers for their detailed feedback.  We're happy the
reviewers found the paper "a joy to read", "an excellent demonstration 
of occurrence typing's power", and provided "substantial practical 
contribution".

We begin with three high-level issues:

1. What are the contributions beyond occurrence typing as presented
   in ICFP'10?

As described in Section 1, our overall contributions are

  - a type system for Clojure supporting Clojure idioms,
  - a sound formal model for three crucial type system features,
  - an implementation successfully used in practice and industry, and
  - an evaluation of the type system in practice.

There are two main extensions specifically related to occurrence typing.

The first is the presentation of HMap's, with support for flow
checking on sub-parts of maps. This required several extra typing
rules (Fig. 10) and subtle extensions to the `update` rule (Fig. 12,
clauses 4-7).

The second is multimethods, which is mainly contained in the
T-DefMethod typing rule. It reuses the machinery T-If, which was
present in ICFP'10, but demonstrates occurrence typing is able to
support new and previously-unhandled control flow.

2. Are multimethods worth the extra complexity in the type system, if
   they are used rarely?

Yes.

The numbers presented in the evaluation section do not indicate that
multimethods are are.  Even though a low number of distinct
multimethods are defined in practice, extending them via `defmethod`
accounts for a large amount of code.

Analagously, though Java code may define few interfaces, they will be
used in more class definitions, and then in many method calls.

As noted in Section 5.1, these multimethods are often single-dispatch
on a fixed map entry, making them even more similar to Java
interfaces, which perform single dispatch on the first argument of
methods.

3. What is the distinction between optional and gradual typing?

Optional typing refers to type systems that do not enforce type invariants
at runtime, and are comparable to Bracha's "pluggable" type systems (as
suggested by Review 2).

Typed Clojure does not check assumptions at runtime, thus is not
gradually typed, just optionally typed like TypeScript.

Review 2 requests to categorize the following:
 - C# 4.0: gradually typed. 
 - Scala's Dynamic: a mostly-unrelated metaprogramming construct, 
   comparable to a statically resolved "method_missing" in Ruby. 
 - Dart: optionally typed.

A gradual type system is defined by its operational semantics, in particular
that the runtime prevents untyped code from breaking static invariants. An
addition of an "Unknown or Dynamic type" (Review 2) is not enough to make
a gradual type system without this enforcement.

We distinguish optional and gradual typing in footnote 1; see [Siek et
al, SNAPL'15] for more on the distinction.

"Any" is the top type, effectively the same as `(U nil Object)`, and
has no special runtime semantics.  It is *not* a Dynamic type (Review
2).

==========================================================================

We now address a few specific questions from the reviewers. The
reviewers also make several suggestions for presentation revisions and
improvements to the evaluation, including comparisons to other systems---we 
will make these changes in a revised version.

REVIEW 1: Is TC decidable? `restrict` seems problematic.

  The calculus leaves the details of `restrict` up to
  implementations. core.typed features an ad-hoc intersection type, so
  `restrict` simply returns the intersection of both types, thus
  making this function simple.

REVIEW 1: Why can't the type system check Example 9? 
          What restrictions are imposed on HMaps?

  For presentational reasons we only support a specific pattern of
  looking up a common, known key in a union of HMaps (notice T-GetHMap
  (Fig. 10) is the only rule supporting lookups on unions of
  HMaps). Example 9 is checkable in our implementation, but not the
  calculus, and serves as a demonstration of the kinds of patterns
  that might occur in Clojure code.

REVIEW 1: What specific patterns could not be supported by TC that led
          to CircleCI abandoning it?

  We refer to the detailed blog post they published:

    http://blog.circleci.com/why-were-no-longer-using-core-typed/

REVIEW 3: Are the case studies representative programs that would be
          good candidates for optional typing?

  We chose the CircleCI code base as it was a substantial size. The
  other case study was a randomly chosen open source project. A goal
  of Typed Clojure is to support as many Clojure idioms as possible,
  and we learnt valuable lessons from both projects in this regard.

REVIEW 1: How is TC compiled/used in practice? Is it sealed from
untyped namespaces?

  TC is compiled to as normal Clojure code. As discussed above, Typed
  Clojure is not gradually typed.

REVIEW 1: "Java cannot be passed nil"?

  This is a conservative assumption, as we mention.  In practice this
  is a reasonable default, as discussed in the CircleCI evaluation
  (5.1).

REVIEW 1: Does (dissoc k) just remove its argument-entry from a map?

  (dissoc m k) returns a new map that is like m, but without a k entry.

REVIEW 1: How do symbolic representations of objects in OT relate to Scala's
          path-dependent types?

  They are unrelated.

REVIEW 1: Formula (3) -- do the positive/negative statements mean that d
          is of type Nat, so N_d should be read as a proposition?

  Yes, N_d says d is a Nat, \overline{N_d} says d is not a Nat.

REVIEW 1: Are Java's generic classes (in particular collections) supported
          in TC type checking algorithm?

  No. TC however supports parametric polymorphism for Clojure
  collections and a manual subset of Java collections. In practice,
  this is sufficient for most interop situations.

REVIEW 1: The simplified syntax chosen for method calls works for
          compilers, but not IDE's. How do Clojure IDE's
          "incrementalize" type checking for code completion?

  This is not our choice of syntax; (.method o) is sugar for (. o
  (method)) in Clojure.  The Cursive Clojure IDE prioritizes type
  hints assigned to variables in the local scope during method name
  autocompletion.

REVIEW 1: "Then (new[S] F "a/b") is checked against a nilable conversion..."?

  The prose should be fixed to instead read "non-nilable".

REVIEW 1: What is the type of (defmethod e_m e_v e_f)?

  As shown in T-DefMethod, it returns the same type as e_m.

REVIEW 1: How are overlapping multi-method clauses handled?

  In the type system, each method is checked in isolation with respect
  to the current dispatch value and the dispatch function type.  In
  the operational semantics, B-BetaMulti can only fire if there is an
  unambiguous match, encoded in the GM metafunction.

REVIEW 2: What is "ann"?

  (ann v t) 'ann'otates variable v as type t. We will explain this
  explicitly.

REVIEW 2: What is the "Kw" type?

   "Kw" is the type for keywords.

REVIEW 2: Why does 42 have the object "0"?

  Objects only describe references to the runtime environment, so we
  say the object of 42 is "empty".