### 2018-12-09 small.tex

% rubber: set arguments --shell-escape
% arara: pdflatex: {shell: true}
%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass{llncs}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\bibliographystyle{splncs03}

\usepackage{common}
\include{bibliography.bib}
\usepackage{chngcntr}
\usepackage{url}
% seems like a nightmare to use, but you get clickable links
%\usepackage{hyperref}

\newcounter{ex}
\newenvironment{exmp}{
\refstepcounter{ex}%
\par\vspace{.8em}\hfill\framebox{\small Example~\theex}\nopagebreak
\\[-3.23em]\nopagebreak}
{\relax}

\newenvironment{exmp2}{
\refstepcounter{ex}%
\par\vspace{3em}\hfill\framebox{\small Example~\theex}\nopagebreak
\\[-3.23em]\nopagebreak}
{\relax}

\newenvironment{mintedclj}{
\begin{verbatim}%
}
{\end{verbatim}%
}%

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

%
%\setminted{%fontsize=\footnotesize
%}

\newcommand{\clj}[1]{\lstinline{#1}}
\newcommand{\java}[1]{\lstinline{#1}}
\newcommand{\rkt}[1]{\lstinline{#1}}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

%\titlebanner{banner above paper title}        % These are ignored unless
%\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Practical Optional Types for Clojure}

%\authorinfo{}{}{\vspace*{-2mm}}
\author{\ma{\text{Ambrose Bonnaire-Sergeant}^\dagger\text{, Rowan Davies*}, \text{Sam Tobin-Hochstadt}^\dagger}}
\institute{Indiana University$\dagger$; Omnia Team, Commonwealth Bank of Australia*
  \{abonnair,samth\}@indiana.edu, Rowan.Davies@cba.com.au}
%       {Indiana University}
%       {abonnair@indiana.edu}
%       {University of Western Australia}
%       {Rowan.Davies@cba.com.au}
%       {Indiana University}
%       {samth@indiana.edu}

\maketitle

\begin{abstract}
  \input{abstract}
\end{abstract}

%% \category{CR-number}{subcategory}{third-level}

%% \keywords
%% optional types, Clojure

\input{intro}

\input{overview}

\input{formal-model}

\input{metatheory}

\input{experience}

\input{related-work}

\input{conclusion}

%\input{acknowledgements}

% We recommend abbrvnat bibliography style.

% The bibliography should be embedded for final submission.
\newpage

\bibliography{bibliography}

%\clearpage
%
%\input{appendix}

\end{document}

%                       Revision History
%                       -------- -------
%  Date         Person  Ver.    Change
%  ----         ------  ----    ------

%  2013.06.29   TU      0.1--4  comments on permission/copyright notices